Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-10-05 exp1/present2013.tex
Effects and simpler design matrix
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@648 43ef1190-90fd-4dd7-a769-d0121f02b80e

For example:
\begin{itemize}
\item Hardware platform
\item Data heterogeneity
\item Number of triples in the store
\item Absence or presence of certain language features
\item Order of experiments
\end{itemize}

@@

For example:
\begin{itemize}
\item For number of triples, the levels could be 1 or 2 MTriples.
\item For language features, the levels could be absence or presence
  of an \texttt{OPTIONAL} pattern. 
\end{itemize}

Levels may be continuous, discrete, two different instances of a
class, etc.

@@

\begin{frame}{Design Matrix}
% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Fri Oct  4 22:47:01 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{r|rrrrrrrr}
  \hline
 & ``Implement'' & ``TripleC'' & ``Machine'' & ``BGPComp'' & ``Lang'' & ``Range'' & ``Union'' & ``Optional'' \\ 
  \hline
1 & 2 & 2 & 1 & 2 & 1 & 2 & 2 & 2 \\ 
  2 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 \\ 
  3 & 1 & 2 & 1 & 2 & 2 & 2 & 1 & 2 \\ 
  4 & 1 & 1 & 2 & 1 & 2 & 2 & 1 & 1 \\ 
  5 & 1 & 1 & 1 & 2 & 1 & 2 & 2 & 2 \\ 
  6 & 1 & 2 & 1 & 2 & 2 & 1 & 1 & 1 \\ 
  7 & 1 & 1 & 2 & 2 & 1 & 2 & 2 & 1 \\ 
  8 & 2 & 1 & 2 & 2 & 2 & 2 & 1 & 2 \\ 
 ...  \\
  256 & 1 & 1 & 2 & 2 & 1 & 1 & 1 & 1 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}
\end{frame}

>>>>

For example:
\begin{itemize}
\item A concrete implementation
\item Hardware platform
\item Data heterogeneity
\item Number of triples in the store
\item Absence or presence of certain language features
\item Order of experiments
\end{itemize}

@@

For example:
\begin{itemize}
\item For implementation, 4store or Virtuoso.
\item For number of triples, the levels could be 1 or 2 MTriples.
\item For language features, the levels could be \textsf{SELECT} and \textsf{CONSTRUCT}
\end{itemize}

Levels may be continuous, discrete, different instances of a
class, etc.

@@

\begin{frame}{Design Matrix}

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Sat Oct  5 22:16:51 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{r|ccc}
  \hline
 & Implement & TripleC & Union \\ 
  \hline
1 & 2 & 1 & 2 \\ 
  2 & 2 & 2 & 2 \\ 
  3 & 2 & 1 & 1 \\ 
  4 & 1 & 1 & 1 \\ 
  5 & 2 & 2 & 1 \\ 
  6 & 1 & 2 & 1 \\ 
  7 & 1 & 2 & 2 \\ 
  8 & 1 & 1 & 2 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

\end{frame}
\note{There are 3 factors, 8 experiments, which is $2^3$}

\begin{frame}{Factorial experiments}

  \begin{itemize}
  \item Full factorial experiments:
    \begin{itemize}
    \item 2 levels give $2^n$ combinations (\emph{runs}), where $n$ is
      the number of factors.
    \item 3 levels give $3^n$ runs, etc.
    \end{itemize}
  \item Fractional factorial experiments:

    \begin{itemize}
    \item When experimental economy is important,
    \item DoE theory offers extensive facilities for running a
      fraction of the runs ($2^{n-1}$ etc).
    \item The price is of smaller experiments is explanatory power.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Main effects}

\emph{Effects} describe the influence of the levels on the response.

The the \emph{main effect}: For a factor $A$ with two levels, we let
$a_1$ and $a_2$ be the average response of all $2^{n-1}$ measurements
with $A$ at level 1, and 2, respectively.  The main effect of $A$ is
then defined as $a_2 - a_1$.

For example:

For each implementation, average all the responses. Then, the main
effect of the implementation is the difference between them.

\begin{frame}{Interaction effects}

\emph{Interaction effects} for two factors $AB$ are defined
by comparing averages for equal versus non-equal levels of $A$ and
$B$.

For example:

For each implementation, find the averages for each dataset size, the
interaction effect is % TODO: Clear example.

In practice, Quite Easily Done using linear regression.
\end{frame}
\note{Interaction effects may be complex, like one implementation is
  better on CONSTRUCT queries, on small datasets}
\note{We still need to understand which effects are important and
  significant, we do that in the discussion of our experiment]
